### 2018-01-29 Physics Notes/Nuclear Physics/Radioactivity/Radioactivity.tex
Physics notes update

\includegraphics[width=6cm]{foil.jpg}  
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
\begin{tabularx}{\textwidth}{|X|X|}
\hline
Observation&Explanation\\
\hline
Most electrons pass all the way through&Atoms are mostly empty space\\
\hline
Some are deflected&The atom has a positive centre\\
\hline
Some are deflected by significant angles&The positive charge is condensed in a small area\\
\hline
\end{tabularx}
    \end{minipage}
\end{figure}
\subsection{Estimating the size of the nucleus}  
\subsubsection{Closest approach method}
\begin{tikzpicture}
\filldraw[color=red, fill=red, very thick](0,0) circle (0.2);
\draw[thick, ->,black] (0.4,0) -- (4,0) node[above,yshift=0.3cm,xshift=-2cm] {\footnotesize{Velocity reduces on approach}};
\draw[color=red, very thick](6,0) circle (1) node[] {Au};
\draw[color=black, very thick,dashed](6,0) circle (2) node[above,yshift=2cm] {EPE};
\draw[thick, ->,black] (4,0) -- (1,-1);
\end{tikzpicture}\\
\\
KE=EPE
$$8.0\times10^{-13}=\frac{1}{4\pi\epsilon_0}\times\frac{Q_{Au}}{r}\times Q_{\alpha}$$
$$r=4.55\times10^{-14}$$
\subsubsection{Estimate from scattering data}
\begin{itemize}
\item About $\frac{1}{10,000}$ deflected through more than 90\degree
\item Foil had n layers of atoms
\end{itemize}
\begin{tikzpicture}
\draw[thick, stealth-stealth,black] (6.8,-0.4) -- (6.8,0.4) node[right,midway] {d} ;
\draw[color=red, very thick](6,0) circle (0.5) node[] {\tiny{Nucleus}};
\draw[color=black, very thick](6,0) circle (2) node[above,yshift=2cm] {Atom};
\draw[thick, stealth-stealth,black] (8.5,-2) -- (8.5,2) node[right,midway] {D};
\end{tikzpicture}\\
$$\frac{\frac{1}{4}\pi d^2}{\frac{1}{4}\pi D^2}=\frac{d^2}{D^2}=\frac{1}{10,000n}$$
$n=10^4$ layers
$$\frac{d^2}{D^2}=\frac{1}{10,000\times1\times10^4}$$
$$d=\frac{D}{10,000}$$
\end{document}

>>>>

\includegraphics[width=6cm]{foil.jpg}  
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
\begin{tabularx}{\textwidth}{|X|X|}
\hline
Observation&Explanation\\
\hline
Most electrons pass all the way through&Atoms are mostly empty space\\
\hline
Some are deflected&The atom has a positive centre\\
\hline
Some are deflected by significant angles&The positive charge is condensed in a small area\\
\hline
\end{tabularx}
    \end{minipage}
\end{figure}
\subsection{Estimating the size of the nucleus}  
\subsubsection{Closest approach method}
\begin{tikzpicture}
\filldraw[color=red, fill=red, very thick](0,0) circle (0.2);
\draw[thick, ->,black] (0.4,0) -- (4,0) node[above,yshift=0.3cm,xshift=-2cm] {\footnotesize{Velocity reduces on approach}};
\draw[color=red, very thick](6,0) circle (1) node[] {Au};
\draw[color=black, very thick,dashed](6,0) circle (2) node[above,yshift=2cm] {EPE};
\draw[thick, ->,black] (4,0) -- (1,-1);
\end{tikzpicture}\\
\\
KE=EPE
$$8.0\times10^{-13}=\frac{1}{4\pi\epsilon_0}\times\frac{Q_{Au}}{r}\times Q_{\alpha}$$
$$r=4.55\times10^{-14}$$
\subsubsection{Estimate from scattering data}
\begin{itemize}
\item About $\frac{1}{10,000}$ deflected through more than 90\degree
\item Foil had n layers of atoms
\end{itemize}
\begin{tikzpicture}
\draw[thick, stealth-stealth,black] (6.8,-0.4) -- (6.8,0.4) node[right,midway] {d} ;
\draw[color=red, very thick](6,0) circle (0.5) node[] {\tiny{Nucleus}};
\draw[color=black, very thick](6,0) circle (2) node[above,yshift=2cm] {Atom};
\draw[thick, stealth-stealth,black] (8.5,-2) -- (8.5,2) node[right,midway] {D};
\end{tikzpicture}\\
$$\frac{\frac{1}{4}\pi d^2}{\frac{1}{4}\pi D^2}=\frac{d^2}{D^2}=\frac{1}{10,000n}$$
$n=10^4$ layers
$$\frac{d^2}{D^2}=\frac{1}{10,000\times1\times10^4}$$
$$d=\frac{D}{10,000}$$
\newpage
\section{Radioactive materials}
\subsection{Sources of background radiation by most common}
\begin{enumerate}
\item Air (e.g. radon gas)
\item Medical
\item Ground and buildings
\item Food and drink
\item Cosmic rays
\item Nuclear weapons
\item Air travel
\item Nuclear power
\end{enumerate}
\subsection{Geiger M\"{u}ller tube}
\includegraphics[width=10cm]{geiger.png}\\
When a particle of ionising radiation enters the tube, the particle ionises the gas atoms along its track. The negative ions are attracted to the rod and the positive ions to the wall. These ions cause further ionisation, creating enough ions for a current to flow. A pulse of charge passes round the circuit through resistor R, causing the voltage pulse across R which is recorded as a single count by the pulse counter\\
\\
The dead time of the tube, the time taken to regain its non conducting state after an ionising particle enters it, is typically of the order of 0.2ms.
\newpage
\begin{tabularx}{\textwidth}{|c|X|X|X|}
\hline
&Alpha&Beta&Gamma\\
\hline
Nature&2 Protons+2 Neutrons&High speed electron or positron&High energy photon\\
\hline
Range&Up to 10cm&Up to 1m&Infinite\\
\hline 
Deflection in a magnetic field&Deflected&Opposite direction to $\alpha$ particles and more easily deflected&Not deflected\\
\hline
Absorption&Paper&Aluminium&Lead \\
\hline
Ionisation&$10^4$ ions per mm&100 ions per mm&Very weak ionising effect\\
\hline
Energy of each particle&Constant for a given source&Varies up to a maximum for a given source&Constant for a given source\\
\hline
\end{tabularx}

\end{document}
